---
title: "Data Tidying Lesson"
author: "Timothy Pasch"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr) 
library(tidyr)
```


stats::filter
```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
```

```{r}
catch_data1 <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
```


```{r}
summary(catch_data)
```

```{r}
  catch_clean <- catch_data %>%
  mutate(Chinook =as.numeric(Chinook))
```

```{r}
i <- which(is.na(catch_clean$Chinook))
```

```{r}
catch_clean <- catch_data %>%
mutate(Chinook = ifelse(Chinook %in% c("I", "l"), 1, Chinook))
```

```{r}
catch_long <- catch_clean %>%
  pivot_longer(cols = -c(Region, Year),
names_to = 'species',
values_to = "catch")


```
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to = "catch")
```
```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))
```

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                  stringsAsFactors = FALSE)
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch") %>%
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  left_join(region_defs, by = c("Region" = "code"))

head(mean_region)

```{r}
```

```{r}
mean_region <- catch_long %>%
  filter(species == "Sockeye") %>%
  group_by(Region) %>%
  summarise(mean_catch = mean(catch))

mean_sockeye region
```

```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>%
  summarise(catch_mean = mean(catch))

head(mean_region)
```
```{r}
n_region <- catch_long %>% 
  group_by(Region) %>%
  summarize(n = n())

head(n_region)
```
```{r}
SSE_catch <- catch_long %>% 
  filter(Region == "SSE")

head(SSE_catch)
```
```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(mean_catch)

head(mean_region)
```

```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(desc(mean_catch))

head(mean_region)
```
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_long, region_defs, by =c("Region" = "code"))

head(catch_joined)

```

```{r}
region_defs <- region_defs %>% 
  rename(Region = code, Region_Name = mgmtArea)

catch_joined <- left_join(catch_long, region_defs, by = c("Region"))

head(catch_joined)
```
```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  separate(date, c("month", "day", "year"), "/")

```
```{r}
cities_df <- data.frame(city = c("Juneau AK", 
                                 "Sitka AK", 
                                 "Anchorage AK"),
                        stringsAsFactors = FALSE)
```

```{r}
cities_df <- data.frame(city = c("Juneau AK", 
                                 "Sitka AK", 
                                 "Anchorage AK"),
                        stringsAsFactors = FALSE)
```

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                  stringsAsFactors = FALSE)
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

mean_region <- catch_original %>%
  select(-All, -notesRegCode) %>% 
  mutate(Chinook = ifelse(Chinook == "I", 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook)) %>% 
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch") %>%
  mutate(catch = catch*1000) %>% 
  group_by(Region) %>% 
  summarize(mean_catch = mean(catch)) %>% 
  left_join(region_defs, by = c("Region" = "code"))

head(mean_region)
```

```


